Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2015-10-01 Publications/2015-NeuroImage/responseLetter2.tex

\begin{reviewpoint}
However, in such a case the paper should be thoroughly revised to address the following major issues:
1. The assumption of independence between pixels is quite usual but obviously wrong.
The authors may want to comment on the consequences.
\end{reviewpoint}
\reply{
The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification.
We have supported the assumptions inserting the new lines 80--83.}

\begin{reviewpoint}
2. It is unclear, under which conditions enforcing the optimality condition guarantees that the region identified corresponds
  to the targeted one (it probably strongly depends on its intensity properties and those of neighbouring structures).
A corresponding analysis of the goal function and the presence of (undesired) local minima is missing.
Figure S1 is not particularly helpful in this respect.
Some information about the initialization of the distortion map should also be given.
\end{reviewpoint}
\reply{
We have improved former Figure 4 (now, it is Figure 2) to clarify that the segmentation model is inherently defined by
  the choice of surfaces and their nesting.
This is an actual contribution of the work, as stated in ll. 264--266.

The model behaves robustly when a) the target data are simple as in our phantom settings; and b) when the distortion
  has only one nonzero component like in our real-data settings.
The analysis of the goal function for the phantoms (without directional restrictions) was not necessary since
  convergence was straightforward.
In the case of real data, since the distortion under investigation is aligned with one axis, the Figure S1 shows
  that the goal function is well defined.

@@

\begin{reviewpoint}
5. The vertex-wise error distributions are not only skewed, but quite clearly structured im many cases
  (especially for the phantom images).
What is the underlying reason?
\end{reviewpoint}
\reply{
We have added a discussion on this in the corresponding section, lines 286-291. Thanks for pointing out.
}

\bigskip
\bigskip
\subsection*{Reviewer \#2:}
\begin{reviewpoint}
This paper seems to present a method that can perform wM-GM segmentation and simultaneously estimate a deformation field to
  correct for EPI distortion.
Apparently, the idea is to use nested WM-GM surfaces, as well as other surfaces, to provide better cortical registration than
  a competing method, which is not clearly described.
The overall goal and motivation of the paper are not clearly explained, and almost all related references for EPI distortion
  correction are missing.
It is unclear if the deformation field solution is restricted to the phase encode direction.
It is possible that this is a very useful method, as the results seem to indicate, but the paper is unfortunately too unclear to
  make a full judgement about the method.
\end{reviewpoint}
\reply{
We thank the reviewer for pointing out that the overall goal is not clearly explained.
We have replaced the term ``active-contours'' by ``surfaces'' whenever possible.
We have re-organized the introduction fully, as suggested (see reply to \autoref{com:1} for a brief explanation).
We have included references of the main EPI distortion corrections (see lines 31-43).

The \emph{regseg} tool is designed to allow any displacement, as demonstrated with the phantoms.
However, as the reviewer points out, the deformation field is completely restricted to the phase-encode direction
  for the real datasets.
We have added ll. 125--127 to indicate that the ground-truth in the experiments done with real data is directionally
  restricted. Then, in lines 190-192 to clarify that \emph{regseg} is configured to fulfill this feature.
}

\begin{reviewpoint}
abstract: Not immediately clear what this means: ``active contours without edges extracted from structural images''
  (Where do the edges come from? What modalities are actually input to regseg?)
\end{reviewpoint}
\reply{%
In our adaptation of ``active contours without edges'' \citep{chan_active_2001}, the active contours are
  ``extracted from structural images''.
Then, they are ``without edges'' because the target features do not show clear intensity steps
  where the contours are intended to be finally placed.
In practice, active contours are ``without edges'' when they look for homogeneous regions enclosed
  by the contours, as is the case of the FA and ADC.
We agree with the reviewer on that this phrase is not clear, and therefore reformulated the abstract.
We have also modified the introduction to state more clearly the actual modalities that input to \emph{regseg}.
}

@@

[...]

I still see nothing in the results indicating that the estimated distortions were restricted to the phase encode direction.
If not, the added degrees of freedom might result in improved surface segmentation, but the distortion field would not correspond at
  all to the physical problem under study.
Furthermore, the dMRI gradients would have to be re-oriented independently at each voxel, and this is not currently possible in any
  software I know of.
\end{reviewpoint}
\reply{
The displacements field supporting the deformation model is not restricted to the phase encoding direction by default.
Therefore, equations in Methods section are presented under such conditions.
The proof of concept presented in the first experiment with phantoms are conducted without preferential directions
  for the deformation.
However, the point raised about the distortions derived from susceptibility inhomogeneity in dMRI data is also true.
Therefore, experiments with real data mimicking such distortions are performed with the deformation aligned with the phase encoding
  axis.
New lines 203-205 now state this point clearly.
}

\begin{reviewpoint}
The experiments seem reasonable and clearly presented in general. But far too little detail is given on the B0-T2 comparison method.
\end{reviewpoint}
\reply{
We thank the reviewer for positively weighing our experiments and results.
As stated in line 214, we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
  internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
A footnote in line 215 directly links the settings for this method.
}

\begin{reviewpoint}
Please define the phrase ``active contours without edges''.
\end{reviewpoint}
\reply{We hope this point is answered in \autoref{com:10}, and tried to avoid this concept
  coming from the 2D processing field.
Now, we refer to the contours as surfaces whenever possible.
Regarding the ``without edges'' feature, we have clarified how the surfaces enclose homogeneous
  regions (and therefore, they are not used to search for image edges).
}

### 2015-10-01 Publications/2015-NeuroImage/02-abstract.tex
fixed typos, added comment labels in manuscript

% -*- root: 00-main.tex -*-
\newcomment[RV\#1(C.1)]{%
Current methods using diffusion MRI to map the microstructure and connectivity of the
  human brain increasingly require precise delineations of anatomical structures.
The problem is typically solved by projecting the structural information extracted
  in T1-weighted images, by means of nonlinear registration since diffusion
  MRI data present geometrical distortions.
Here we propose \regseg{}, which segments homogeneous regions within multivariate
  images by projecting a set of nested surfaces extracted from structural images.}
The surfaces evolve through a free-form deformation field supported by the B-spline basis
  to optimally map the contours onto the data in the target space.
We tested the functionality of \regseg{} using four digital phantoms warped with known and
  randomly generated deformations, where subvoxel accuracy was achieved.
\newcomment[RV\#2(C.11)]{%
We proposed a set of surfaces that define a segmentation model of the fractional anisotropy and
  the apparent diffusion coefficient maps derived from diffusion MRI datasets.}
These surfaces extracted from the same-subject T1-weighted image successfully segmented 16 real
  diffusion MRI datasets from the Human Connectome Project that were warped by realistic and nonlinear
  distortions.
We computed the misregistration error of the surfaces estimated by \regseg{} with respect
  to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
  distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
We also compared the performance of the proposed method with a widely used registration tool, which showed
  that \regseg{} outperformed this method in our settings.

>>>>

% -*- root: 00-main.tex -*-
\newcomment[RV\#1(C.1)\par RV\#2(C.10)]{%
Current methods using diffusion MRI to map the microstructure and connectivity of the
  human brain increasingly require precise delineations of anatomical structures.
The problem is typically solved by projecting the structural information extracted
  in T1-weighted images, by means of nonlinear registration since diffusion
  MRI data present geometrical distortions.
Here we propose \regseg{}, which segments homogeneous regions within multivariate
  images by projecting a set of nested surfaces extracted from structural images.}
The surfaces evolve through a free-form deformation field supported by the B-spline basis
  to optimally map the contours onto the data in the target space.
We tested the functionality of \regseg{} using four digital phantoms warped with known and
  randomly generated deformations, where subvoxel accuracy was achieved.
\newcomment[RV\#2(C.11)]{%
We proposed a set of surfaces that define a segmentation model of the fractional anisotropy and
  the apparent diffusion coefficient maps derived from diffusion MRI datasets.}
These surfaces extracted from the same-subject T1-weighted image successfully segmented 16 real
  diffusion MRI datasets from the Human Connectome Project that were warped by realistic and nonlinear
  distortions.
We computed the misregistration error of the surfaces estimated by \regseg{} with respect
  to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
  distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
We also compared the performance of the proposed method with a widely used registration tool, which showed
  that \regseg{} outperformed this method in our settings.
